Considering scale in multilayer animal social networks

### 2020-06-14 paper/manuscript.Rmd
tidy

## Caribou socioecology
<!--Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on predicted social (visual and auditory perception), spatial
(habitat type), and temporal (seasonal) scales.--> Caribou are gregarious ungulates
with fission-fusion dynamics [@Lesmerises_2018] that display temporal and
spatial variation in social networks [@Peignier_2019]. In winter, caribou dig
holes in the snow, termed craters, to access forage [@Bergerud_1974]. Cratering
is presumed to be a costly behaviour and as a result, caribou tend to occupy and
re-use craters once they are established. <!-- and there is considerably less access
to forage than when the landscape is snow-free.--> Importantly, craters exist on
the landscape at multiple scales: the crater scale (a single crater), the
feeding area scale (multiple craters in close proximity), and the winter range
scale (all craters within an individual’s range) [@Mayor_2009]. Access to forage is 
seasonally heterogeneous depending on snow cover and, in winter, variable distribution of craters on the 
landscape. <!--We partitioned data into two
discrete seasons: summer (3 August - 20 September 2017) and winter (1 January – 17
February 2018) based on caribou socioecology [@Peignier_2019].-->

@@

We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r version$version.string` [@R_Core_Team_2019]. We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the spatial distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a spatial distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018].

<!--In
all networks, individual caribou were represented as nodes, associations between
individuals in a given habitat type and season were represented as intralayer
edges, and connections between the same individuals across seasons and within
habitat classes were represented as interlayer edges (Figure \@ref(fig:ml)) .-->
<!-- eg this is connecting contexts used to calculate layer similarity -->

@@

### Network metrics 
<!-- TODO: expand + simplify + examples individual A, B etc -->

We used a series of metrics to characterize the multilayer networks, focusing on
the role of individuals within and across layers, and the similarity of
different layers. To measure the role of individuals within layers, we
calculated degree centrality and graph strength. Degree centrality is the number
of direct connections an individual has to other individuals in a network layer.
Graph strength is the degree weighted by the strength of association, in this
case SRI, in each layer. Extending degree centrality, we calculated multidegree,
the sum of degree centrality of individuals across layers  [@Berlingerio_2012;
@Kivela_2014]. <!--We evaluated the relevance and redundancy of layers by
combining the multidegree and neighbourhood to calculate layer relevance and
connective redundancy. Layer relevance is the proportion of neighbours present
in each layer compared to the neighbourhood for each individual and connective
redundancy is the proportion of neighbours repeated in multiple layers, and
[@Berlingerio_2012].--> Finally, we measured the similarity of layers by
calculating edge overlap. Edge overlap is defined as the proportion of edges
present in each network layer out of all observed edges  [@Battiston_2014].
<!--between seasons, across land cover classes (e.g. summer open compared to
winter open) using the Pearson correlation coefficient [@Brodka_2018].-->
Network metrics were calculated using the R packages `igraph` [@Csardi_2006],
`asnipe` [@Farine_2019], `spatsoc` [@Robitaille_2019] and `data.table`
[@Dowle_2019].

## Varying scale in multilayer networks
<!--

social scale: variable is spatial distance threshold
spatial scale: variable is land cover resolution
temporal scale: variable is temporal window
temporal scale: variable is number of observations
-->

@@

### Temporal scale
#### Time windows

We used a temporal multilayer network to assess the seasonality of caribou
sociality. Multilayer networks consisted of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across 20 ordinal sample periods of 35-36
days taken from the entire study period. Social associations were defined by
spatiotemporal overlap of individuals within 5 minutes and 50 m. In each time
window, we calculated individual graph strength and multidegree, and across the
series of networks, we calculated edge overlap to determine proportion of total
edges observed in each time window.

@@

# Results
<!-- TODO: edge overlap can indicate if some associations are dependent on context -->
<!-- TODO: Add intro/general to results?-->
<!-- major points then details -->

@@

Increasing land cover resolution resulted in decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats at higher resolution (Figure \@ref(fig:lcres)). Graph strength
and edge overlap were relatively consistent across land cover resolutions for
forest and open layers (Figure \@ref(fig:lcres) D, E). Edge overlap decreased
sharply between 30 m and 600 m in forage layers, afterwards remaining stable to
1000 m (Figure \@ref(fig:lcres) D). Graph strength was more variable within
forage layers across land cover resolutions than open and forest (Figure
\@ref(fig:lcres) D). The proportion of relocations in forest and open habitats
increased with increasing spatial resolution, while decreasing in forage habitat
(Supp. 1 <!-- todo-->).  Overall, these results indicate the importance of
matching land cover resolution to scale of selection as well as ensuring the
land cover product used has a resolution fine enough to detect landscape
features and habitats relevant to study species, such as important patchy and
rare habitats.

@@

<!-- TODO: reduce to this version -->
As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics generally appeared to stabilize. Graph
strength across individuals in all three habitat layers was highly variable at low number of observations and stabilized after ~100 observations (Figure \@ref(fig:nobs) A). Edge overlap increased across all three habitat layers with increasing number of observations (Figure \@ref(fig:nobs) B). Similarly, multidegree for all individuals increased quickly between 10 and 100 observations, and continued to increase until 1000 observations (Figure \@ref(fig:nobs) C). Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

<!-- sufficient observations is essential for capturing reliable networ -->

@@

<!-- Figure: metrics by number of observations -->
<!-- TODO: update to temp -->
(ref:nobs) Varying scale in number of observations. Multilayer networks were generated using a  sample number of observations (10 - 1000 observations by steps of 10) of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A) Graph strength was highly variable at low number of observations (less than 100) for all habitat layers and relatively stable after 200 observations. B) Edge overlap increased for all habitat layers with increasing number of observations. Open and forage layers had higher edge overlap than forest layers. C) Multidegree increased for all individuals with increasing number of observations. Individuals showed high variability (`r nobs[nobs == max(nobs), paste(range(multideg), collapse = ' - ')]`) at the maximal number of observations (1000).

@@

<!-- TODO: a bit more for g strength? -->

# Discussion

>>>>

## Caribou socioecology
<!--Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on predicted social (visual and auditory perception), spatial
(habitat type), and temporal (seasonal) scales.--> Caribou are gregarious
ungulates with fission-fusion dynamics [@Lesmerises_2018] that display temporal
and spatial variation in social networks [@Peignier_2019]. In winter, caribou
dig holes in the snow, termed craters, to access forage [@Bergerud_1974].
Cratering is presumed to be a costly behaviour and as a result, caribou tend to
occupy and re-use craters once they are established. <!-- and there is
considerably less access to forage than when the landscape is snow-free.-->
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is seasonally heterogeneous depending on snow
cover and, in winter, variable distribution of craters on the landscape.

@@

We generated proximity-based social networks using the R package `spatsoc`
[@Robitaille_2019] in `r version$version.string` [@R_Core_Team_2019]. We assumed
individuals were associating if simultaneous GPS fixes were within the spatial
distance threshold of one another. Typically for ungulates and other gregarious
mammals, the ‘chain rule’ is used for group assignment [@Croft_2008]. For GPS
data, the chain rule is applied by assigning a group identifier to the union of
buffered GPS fixes at each time step. As such, individuals in a group are within
the spatial distance threshold of at least one other though not necessarily all
other individuals [@Robitaille_2019]. Group assignment based on the chain rule
has previously been applied to caribou using a temporal threshold of 5 minutes
and a spatial distance threshold of 50 m [@Peignier_2019; @Lesmerises_2018]. In
all networks, individual caribou were represented as nodes and associations between
individuals in a given network layer were represented as intralayer
edges. <!-- (Figure \@ref(fig:ml)) .-->

@@

### Network metrics 
We used a series of metrics to characterize the multilayer networks, focusing on
the role of individuals within and across layers, and the similarity of
different layers. To measure the role of individuals within layers, we
calculated degree centrality and graph strength. Degree centrality is the number
of direct connections an individual has to other individuals in a network layer.
Graph strength is the degree weighted by the strength of association, in this
case SRI, in each layer. Extending degree centrality, we calculated multidegree,
the sum of degree centrality of individuals across layers  [@Berlingerio_2012;
@Kivela_2014]. Finally, we measured the similarity of layers by calculating edge
overlap. Edge overlap is defined as the proportion of edges present in each
network layer out of all observed edges  [@Battiston_2014]. Network metrics were
calculated using the R packages `igraph` [@Csardi_2006], `asnipe`
[@Farine_2019], `spatsoc` [@Robitaille_2019] and `data.table` [@Dowle_2019].

## Varying scale in multilayer networks

@@

### Temporal scale
#### Time windows 
We used a temporal multilayer network to assess the seasonality of caribou
sociality. Multilayer networks consisted of the social association between 
`r DT[, uniqueN(get(idcol))]` individuals across 20 ordinal sample periods of 35-36
days taken from the entire study period. Social associations were defined by
spatiotemporal overlap of individuals within 5 minutes and 50 m. In each time
window, we calculated individual graph strength and multidegree, and across the
series of networks, we calculated edge overlap to determine proportion of total
edges observed in each time window.

@@

# Results
<!-- TODO: Add intro/general to results?-->

@@

Increasing land cover resolution resulted in decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats at higher resolution (Figure \@ref(fig:lcres)). Graph strength
and edge overlap were relatively consistent across land cover resolutions for
forest and open layers (Figure \@ref(fig:lcres) D, E). Edge overlap decreased
sharply between 30 m and 600 m in forage layers, afterwards remaining stable to
1000 m (Figure \@ref(fig:lcres) D). Graph strength was more variable within
forage layers across land cover resolutions than open and forest (Figure
\@ref(fig:lcres) D). The proportion of relocations in forest and open habitats
increased with increasing spatial resolution, while decreasing in forage habitat
(Supp. 1 <!-- TODO-->).  Overall, these results indicate the importance of
matching land cover resolution to scale of selection as well as ensuring the
land cover product used has a resolution fine enough to detect landscape
features and habitats relevant to study species, such as important patchy and
rare habitats.

@@

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics generally appeared to stabilize.
Graph strength across individuals in all three habitat layers was highly
variable at low number of observations and stabilized after ~100 observations
(Figure \@ref(fig:nobs) A). Edge overlap increased across all three habitat
layers with increasing number of observations (Figure \@ref(fig:nobs) B).
Similarly, multidegree for all individuals increased quickly between 10 and 100
observations, and continued to increase until 1000 observations (Figure
\@ref(fig:nobs) C). Given the observed trend of these multilayer metrics
stabilizing after a certain number of observations are included, the results
suggest this sensitivity method could be useful for determining the number of
observations necessary for sufficiently describing social dynamics across
contexts.

@@

<!-- Figure: metrics by number of observations -->
(ref:nobs) Varying scale in number of observations. Multilayer networks were generated using a  sample number of observations (10 - 1000 observations by steps of 10) of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland between `r DT[, format(min(datetime), '%F')]` and `r DT[, format(max(datetime), '%F')]` across three habitat classes (forage, forest, and open). A) Graph strength was highly variable at low number of observations (less than 100) for all habitat layers and relatively stable after 200 observations. B) Edge overlap increased for all habitat layers with increasing number of observations. Open and forage layers had higher edge overlap than forest layers. C) Multidegree increased for all individuals with increasing number of observations. Individuals showed high variability (`r nobs[nobs == max(nobs), paste(range(multideg), collapse = ' - ')]`) at the maximal number of observations (1000).

@@

# Discussion
<!-- TODO: edge overlap can indicate if some associations are dependent on context -->
<!-- sufficient observations is essential for capturing reliable networ -->
